# Calculating FAR & RR

After estimating your factual and counterfactual probabilities ($p_1$ and $p_0$), you can now calculate the two primary metrics used to quantify the human influence on an extreme event.

## Fraction of Attributable Risk (FAR)

The **FAR** estimates the proportion of the observed event's risk that is linked to the causal factor being studied.

1.  Use the following equation to calculate the FAR:

$$FAR = 1 - \frac{p_0}{p_1}$$

2.  Interpret the Output:
    -   A FAR of 0.90 suggests that 90% of the risk of the event was attributable to the causal factor.
    -   If $p_1$ is much larger than $p_0$, the FAR will approach 1, indicating a high level of attribution.

## Risk Ratio (RR)

The **RR**, also known as the "likelihood ratio," compares the probability of an event occurring in two different scenarios to determine how much a specific factor increased or decreased the likelihood of that event.

1.  Use the following equation to calculate the RR:

$$RR = \frac{p_1}{p_0}$$

2.  Interpret the Output:

    -   A RR $=$ 2.0 means the event is **twice as likely** to occur in the factual world compared to the counterfactual world.
    -   A RR $=$ 1.0 indicates that the causal factor had **no impact** on the likelihood of the event.
    -   A RR $\le$ 1.0 indicates the causal factor **decreased** the likelihood of the event.
    
::: callout-tip

# Example

Continuing with our synthetic 25°C threshold data, we can calculate these metrics in R using the variables we defined in the Estimation step.

```{r}
#| label: estimate-probabilities
#| include: false

# 1. Generate Synthetic Data (The "Worlds")
set.seed(349)
counterfactual_data <- rnorm(1000, mean = 20, sd = 2) # Cooler "Natural" World
factual_data <- rnorm(1000, mean = 22, sd = 2) # Warmer "Current" World
threshold <- 25

# 2. Define the threshold
u <- 25

# 3. Calculate Probabilities
p1 <- mean(factual_data >= u)
p0 <- mean(counterfactual_data >= u)
```

```{r}
#| label: calculate-metrics
#| echo: true

# 1. Calculate Risk Ratio
risk_ratio <- p1 / p0

# 2. Calculate Fraction of Attributable Risk
far <- 1 - (p0 / p1)

# Display results
cat("Risk Ratio (RR):", round(risk_ratio, 2), "\n")
cat("Fraction of Attributable Risk (FAR):", round(far, 2))
```

**Interpretation: **

-   **RR**: The causal factor has made an extreme heat event above 25°C **`r risk_ratio`** times more likely to occur compared to the natural world.
-   **FAR**: We can conclude that **`r round(far,2)*100`%** of the risk of this heat event was contributed by the causal factor, while only **`r round(1-far,2)*100`%** was due to natural internal variability.
:::